---
title: "Basic Inferential Data Analysis Instructions"
author: "Pribylov V.A."
date: "02 09 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Overview
This is a course project needed to analyze the ToothGrowth data in the R datasets package.

## Loading the data
```{r}
TG <- ToothGrowth
```

## Providing basic summary of the data
```{r}
summary(TG)
OJ <- TG$len[TG$supp == 'OJ']
VC <- TG$len[TG$supp == 'VC']
```

## Comparison of means for two factor-levels of supp column
```{r}
mOJ <- mean(OJ)
vOJ <- var(OJ)
nOJ <- length(OJ)
OJ_confidence_interval <- mOJ + c(-1, 1)*qnorm(0.975)*sqrt(vOJ/nOJ)
paste0('OJ 95% confidence interval is (',round(OJ_confidence_interval[1], 2),
       ', ', round(OJ_confidence_interval[2], 2), ')')
```

```{r}
mVC <- mean(VC)
vVC <- var(VC)
nVC <- length(VC)
VC_confidence_interval <- mVC + c(-1, 1)*qnorm(0.975)*sqrt(vVC/nVC)
paste0('VC 95% confidence interval is (',round(VC_confidence_interval[1], 2),
       ', ', round(VC_confidence_interval[2], 2), ')')
```
As we can see, 95% confidence intervals for OJ and VC are intersecting each other. So let's try find 95% confidence interval for difference of their means.  
```{r}
pooled_variance <- sqrt(vOJ/nOJ + vVC/nVC)
difference_confidence_interval <- mOJ - mVC + c(-1, 1)*qnorm(0.975)*pooled_variance
paste0('95% confidence interval of difference between VC and OJ means is ('
       ,round(difference_confidence_interval[1], 2),', '
       , round(difference_confidence_interval[2], 2), ')')
```
This confidence interval includes 0, so there can be no difference between their means, but for 90% confidence interval mean for OJ is bigger than mean for VC:
```{r}
difference_confidence_interval_90 <- mOJ - mVC + c(-1, 1)*qnorm(0.95)*pooled_variance
paste0('90% confidence interval of difference between VC and OJ means is ('
       ,round(difference_confidence_interval_90[1], 2),', '
       , round(difference_confidence_interval_90[2], 2), ')')
```
## Statistical characteristics for dose column
```{r}
mns <- NULL
vars <- NULL
lens <- NULL
for (i in levels(as.factor(TG$dose))) {
    mns <- c(mns, mean(TG$len[TG$dose == i]))
    vars <- c(vars, var(TG$len[TG$dose == i]))
    lens <- c(lens, length(TG$len[TG$dose == i]))
    
}
```
There are some pairwise comparisons between levels of dose:
```{r}
confidence_interval_05 <- mns[1] + c(-1, 1)*qnorm(0.975)*sqrt(vars[1]/lens[1])
confidence_interval_05
confidence_interval_1 <- mns[2] + c(-1, 1)*qnorm(0.975)*sqrt(vars[2]/lens[2])
confidence_interval_1
confidence_interval_2 <- mns[3] + c(-1, 1)*qnorm(0.975)*sqrt(vars[3]/lens[3])
confidence_interval_2
```
As at the previous case let's find 95% confidence intervals of pairwise differences:
```{r}
difference_confidence_interval_05_1 <- mns[2] - mns[1] + 
    c(-1, 1)*qnorm(0.975)*sqrt(vars[1]/lens[1] + vars[2]/lens[2])
paste0('95% confidence interval of difference between 0.5 and 1 levels means is ('
       ,round(difference_confidence_interval_05_1[1], 2),', '
       , round(difference_confidence_interval_05_1[2], 2), ')')
difference_confidence_interval_1_2 <- mns[3] - mns[2] + 
    c(-1, 1)*qnorm(0.975)*sqrt(vars[2]/lens[2] + vars[3]/lens[3])
paste0('95% confidence interval of difference between 1 and 2 levels means is (' 
       ,round(difference_confidence_interval_1_2[1], 2),', '
       , round(difference_confidence_interval_1_2[2], 2), ')')
difference_confidence_interval_05_2 <- mns[3] - mns[1] + 
    c(-1, 1)*qnorm(0.975)*sqrt(vars[1]/lens[1] + vars[3]/lens[3])
paste0('95% confidence interval of difference between 0.5 and 2 levels means is (' 
       ,round(difference_confidence_interval_05_2[1], 2),', '
       , round(difference_confidence_interval_05_2[2], 2), ')')
```

So with the increasing of dose level, we also can see increasing of mean length.

