---
title: "Statistical Inference Course Project"
author: "Pribylov V.A."
date: "02 09 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview 

This is a course project needed to learn some basics in statistical inference such as Central Limit Theorem, confident intervals, etcetera.

## Variables statement

```{r}
lambda <- 0.2
mns <- NULL
vars <- NULL
```

## Creating the distribution

```{r}
for (i in 1 : 1000) {
    distributions <- rexp(40, lambda)
    mns <- c(mns, mean(distributions))
    vars <- c(vars, var(distributions))
}
```

## Comparison to theoretical values

```{r}
experimental_mean <- mean(mns)
experimental_variance <- mean(vars)

theoretical_variance <- 1/lambda**2
theoretical_mean <- 1/lambda

difference_mean <- theoretical_mean - experimental_mean
difference_variance <- theoretical_variance - experimental_variance
```
So we can see, both mean difference and variance difference between theoretical and simulating values are about to 0:
```{r}
paste0('Mean difference is ', difference_mean)
paste0('Variance difference is ', difference_variance)
```

So we will assume that experimental mean and variance are equal to theoretical using confidence intervals:
```{r}
mean_confidence_interval <- experimental_mean + 
    c(-1, 1)*qnorm(.975)*sqrt(experimental_variance/1000)
variance_confidence_interval <- experimental_variance + 
    c(-1, 1)*qnorm(.975)*sqrt(experimental_variance/1000)
```

```{r}
paste0('Confidence interval for mean is (', mean_confidence_interval[1],
       ', ', mean_confidence_interval[2], ')')
paste0('Confidence interval for variance is (', variance_confidence_interval[1], 
       ', ', variance_confidence_interval[2], ')')
```
And we can see, that experimental mean and variance lies inside of each confidence interval:
```{r}
paste0('Theoretical mean is ', theoretical_mean)
paste0('Theoretical variance is ', theoretical_variance)
```

## Distribution plot

```{r, echo=FALSE}
hist(mns, xlab = 'Means', main = 'Histogram of means')
```

According to this plot, distribution approximates by normal distribution very good.